---
title: "Activity: Street or Road?"
author: "Guilherme Lima"
output: html_notebook
---

## Load in the data

```{r}
#load libraries 
library(tidyverse)
#load in the data here 
# we will only be using the "street-addresses.csv" file from the text
download.file(url="https://mdbeckman.github.io/dcSupplement/data/CMS_ProvidersSimple.rds",
              destfile = "Your-Name-For-The-File.rds")
DataTable <- readRDS("Your-Name-For-The-File.rds")


```

## Recreate Solved Example using `sample(50)`

```{r}
Addresses <- read_csv("https://mdbeckman.github.io/dcSupplement/data/street-addresses.csv")
Sample <- 
  Addresses %>%
  sample_n(size = 50)

SampleMatches <- 
  Sample %>%
  filter(grepl(pattern = "PO", address))

SampleDont <- 
  Sample %>%
  filter( !grepl(pattern = "PO", address))

pattern <- "BOX\\s+(\\d+)"

SampleMatches <- 
  Sample %>% 
  filter(grepl(pattern, address))

SampleDont <-
  Sample %>%
  filter( !grepl(pattern, address))

SampleDont

SampleBoxNumbers <- 
  Sample %>%
  filter(grepl(pattern, address)) %>%
  tidyr::extract(address, into = "boxnum", regex = pattern)
```

## Recreate Soved Example using the entire `Address` data set

```{r}
AddressMatches <- 
  Addresses %>%
  filter(grepl(pattern = "PO", address))

AddressDont <- 
  Addresses %>%
  filter( !grepl(pattern = "PO", address))

pattern <- "BOX\\s+(\\d+)"

AddressMatches <- 
  Addresses %>% 
  filter(grepl(pattern, address))

AddressDont <-
  Addresses %>%
  filter( !grepl(pattern, address))

AddressDont

AddressBoxNumbers <- 
  Addresses %>%
  filter(grepl(pattern, address)) %>%
  tidyr::extract(address, into = "boxnum", regex = pattern)
```

#### Additional Question: How many addresses are a PO BOX (including NCSU BOX)? 

```{r}
AddressBoxNumbers %>% summarise(count = n())

```

My answer here...... 
11100

## Back to the Streets: Your Turn

#### Part 1: Explain (in english) each line of the following code chunk

```{r}
pattern <- "(ST|RD|ROAD)"
#1
LeftOvers <-
#2
  Addresses %>% 
#3
  filter( !grepl(pattern, address),
#4
          !grepl("\\sAPT|UNIT\\s[\\d]+$", address),
#5      
          !grepl(" BOX ", address)
#6    
          )
#7
```

My answer here.......
1) a string indicating an exact pattern match 
2) assign the variable LeftOvers
3) from the Addresses table, use the chaining operator
4) filters addresses for ones that do not match the pattern
5) filters addresses for ones that do not contain APT or UNIT (followed by the number at the end)  
6) filters addresses for ones that do not contain BOX       
7) closes the filter function

#### Part 2: Implement your method on the entire `Addresses` data set. Explain how your code works line by line.

```{r}
pattern <- "(ST|RD|ROAD|CIRCLE|WAY|TRAIL|LN|DR|CIR|AVE|PKWY|COURT|POINT|BLVD|CT|LANE|PL|PATH|LOOP)"
#1
LeftOvers <-   
#2
  Addresses %>%     
#3
  filter( !grepl(pattern, address),    
#4
          !grepl("\\sAPT|UNIT\\s[\\d]+$", address),   
#5
          !grepl("BOX|BX|B0X", address),
#6
          !grepl("HOSPITAL", address))
#7

```

My answer here.......

1) identifies common street endings and applies them to a pattern string
2) assign the variable LeftOvers
3) from the Addresses table, use the chaining operator
4) filters addresses for ones that do not match the pattern
5) filters addresses for ones that do not contain APT or UNIT (followed by the number at the end)
6) filters addresses for ones that do not contain BOX 
7) filters addresses for ones that do not contain HOSPITAL, since hospital is a place and not a street ending

#### Additional Question 1: Present your result from part 2 by providing a table in descending order of popularity for the street name endings you found

```{r}
pattern <- "(ST|RD|ROAD|CIRCLE|WAY|TRAIL|LN|DR|CIR|AVE|PKWY|COURT|POINT|BLVD|CT|LANE|PL|PATH|LOOP)"
StreetEnding <- Addresses %>% filter(grepl(pattern,address)) %>% tidyr::extract(address, into = "ending", regex = pattern) %>% group_by(ending) %>% summarise(count = n()) %>%  arrange(desc(count))
head(StreetEnding)
```

#### Additional Question 2: Present your result by using ggplot to a bar chart in descending order of popularity for the street name endings you found previously.

Note: make sure your order street names according to popularity, not alphabetically.

```{r}
ggplot(StreetEnding, aes(x = reorder(ending,-count), y = count)) +
  geom_bar(stat="identity") + labs(x = "common endings", y= "count of endings", title = "Popularity of street name endings")
```

#### Additional Question 3: Comment on any patterns you see

My answer here.......
